---
title: "Forecasting Inflation"
author: "Abigail Morgan"
date: "`r format(Sys.Date(),'%B %d, %Y')`"
output: 
  html_document:
    df_print: paged
    code_folding: "hide"
    toc: yes
    fig_caption: yes
    theme: spacelab
    toc_float: no
---

```{r setup, include = FALSE}
rm(list=ls())
graphics.off()
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r loadPackages, message=FALSE}
require(fpp3)
require(tidyquant)
require(tidyverse)
require(tsibble)
require(reshape2)
require(kableExtra)
```


## Data
I will be using the following variables to forecast the one-year ahead inflation rate and comparing the accuracy of each forecast model.  Additionally, I will perform a combined forecast using the four specified variables.
```{r loadData, message=FALSE}
# Variables Used
var <- c("PCEPI", "UNRATE", "EXPINF1YR", "INDPRO", "HOUST")

#Pull data from FRED using tidyquant
data <-
  tq_get(var, get = "economic.data", from = "1982-01-01") %>%
  mutate(Month = yearmonth(date), value = price) %>%
  select(-c(date,price)) %>%
  as_tsibble(index = Month, key = symbol)
wdata <- data %>%
  pivot_wider(names_from = symbol, values_from = value) %>% 
  as_tsibble()

# Transform variables so that they are stationary
tdata <- wdata %>% select(c(PCEPI, UNRATE, EXPINF1YR, INDPRO, HOUST))%>%
  mutate(infl = 1200*log(PCEPI/lag(PCEPI))) %>% 
  mutate(dinfl = infl - lag(infl,1)) %>% 
  mutate(dinfl12 = 100*log(PCEPI/lag(PCEPI,12)) - lag(infl,12)) %>% 
  mutate(unrate = UNRATE - lag(UNRATE)) %>%
  mutate(expinf1yr = difference(EXPINF1YR)) %>% 
  mutate(indpro = 1200*log(INDPRO/lag(INDPRO))) %>% 
  mutate(houst = 100*log(HOUST/lag(HOUST))) %>% 
  select(-c(PCEPI, UNRATE, EXPINF1YR, INDPRO, HOUST)) %>% 
  drop_na()
  
# Generate training and testing data sets
train_data <- tdata %>% filter_index(~ "2018-12")
test_data <- tdata %>% filter_index("2019-01" ~ .)
```
 | Variable | Description |  |
|:-----:|:------|----|
| UNRATE |  Unemployment Rate |  |
| EXPINF1YR | 1-Year Expected Inflation | |
| INDPRO |  Industrial Production: Total Index |   |
| HOUST | 	New Privately-Owned Housing Units Started: Total Units | |

```{r plotData, message=FALSE}
# Plot the raw data for each variable
# dataplot <- melt(tdata, "Month")
# ggplot(dataplot, aes(Month, value)) +
#   geom_line() + 
#   facet_wrap(~variable, scales = "free", ncol = 2)
```


## Phillips Curve Model
In order to predict the relationship between unemployment and the inflation rate, I estimated a Phillips curve and used the same specification to estimate the relationship between my other variables of interest and the rate of inflation.
```{r phillipsCurve}
# Unemployment Rate
fitPC <- train_data %>%
  model(
    mU = TSLM(
      dinfl12 ~ 1 +
        lag(dinfl, 12) + lag(dinfl, 13) + lag(dinfl, 14) +
        lag(dinfl, 15) + lag(dinfl, 16) + lag(dinfl, 17) +
        lag(dinfl, 18) + lag(dinfl, 19) + lag(dinfl, 20) +
        lag(dinfl, 21) + lag(dinfl, 22) + lag(dinfl, 23) +
        lag(unrate, 12) + lag(unrate, 13) + lag(unrate, 14) +
        lag(unrate, 15) + lag(unrate, 16) + lag(unrate, 17) +
        lag(unrate, 18) + lag(unrate, 19) + lag(unrate, 20) +
        lag(unrate, 21) + lag(unrate, 22) + lag(unrate, 23)
    ),
# Expected Inflation
    mE = TSLM(
      dinfl12 ~ 1 +
        lag(dinfl, 12) + lag(dinfl, 13) + lag(dinfl, 14) +
        lag(dinfl, 15) + lag(dinfl, 16) + lag(dinfl, 17) +
        lag(dinfl, 18) + lag(dinfl, 19) + lag(dinfl, 20) +
        lag(dinfl, 21) + lag(dinfl, 22) + lag(dinfl, 23) +
        lag(expinf1yr, 12) + lag(expinf1yr, 13) + lag(expinf1yr, 14) +
        lag(expinf1yr, 15) + lag(expinf1yr, 16) + lag(expinf1yr, 17) +
        lag(expinf1yr, 18) + lag(expinf1yr, 19) + lag(expinf1yr, 20) +
        lag(expinf1yr, 21) + lag(expinf1yr, 22) + lag(expinf1yr, 23)
    ),
# Industrial Production
    mI = TSLM(
      dinfl12 ~ 1 +
        lag(dinfl, 12) + lag(dinfl, 13) + lag(dinfl, 14) +
        lag(dinfl, 15) + lag(dinfl, 16) + lag(dinfl, 17) +
        lag(dinfl, 18) + lag(dinfl, 19) + lag(dinfl, 20) +
        lag(dinfl, 21) + lag(dinfl, 22) + lag(dinfl, 23) +
        lag(indpro, 12) + lag(indpro, 13) + lag(indpro, 14) +
        lag(indpro, 15) + lag(indpro, 16) + lag(indpro, 17) +
        lag(indpro, 18) + lag(indpro, 19) + lag(indpro, 20) +
        lag(indpro, 21) + lag(indpro, 22) + lag(indpro, 23)
    ),
# Housing Units
    mH = TSLM(
      dinfl12 ~ 1 +
        lag(dinfl, 12) + lag(dinfl, 13) + lag(dinfl, 14) +
        lag(dinfl, 15) + lag(dinfl, 16) + lag(dinfl, 17) +
        lag(dinfl, 18) + lag(dinfl, 19) + lag(dinfl, 20) +
        lag(dinfl, 21) + lag(dinfl, 22) + lag(dinfl, 23) +
        lag(houst, 12) + lag(houst, 13) + lag(houst, 14) +
        lag(houst, 15) + lag(houst, 16) + lag(houst, 17) +
        lag(houst, 18) + lag(houst, 19) + lag(houst, 20) +
        lag(houst, 21) + lag(houst, 22) + lag(houst, 23)
    )
  )

# report(fitPC)
```


## Forecast Model
Next, I generated a combined forecast model of the three year ahead inflation rate using my four variables of interest.  Looking at the graph, it appears that using the combined model provides the best fitting forecast.
```{r forecast}
fitcombo <- fitPC %>% mutate(combo = (mU + mE + mI + mH )/4)

fc <- fitcombo %>% forecast(new_data = test_data)
fc %>% autoplot(filter(tdata, year(Month) > 2016), level = c(95))
```


## Forecast Accuracy
To verify that the combined forecast was the most accurate, I compared the mean absolute percent errors of the five specifications.  I found that the combined forecast model was in fact the best predictor based on my test data sets results.
```{r accuracy}
 accuracy(fitcombo) %>%
   select(c(".model", ".type", "MAPE"))
 accuracy(fc, tdata) %>%
   select(c(".model", ".type", "MAPE"))
```
